Evaluating Regularization

### 2019-03-12 Scripts/evaluating_feature_reduction.Rmd
updated script for created heatmaps for enet coefficients

@@

>>>>

# Evaluating Elastic-Net

## Heatmaps of whether or not a variable was reduced

@@

## Heatmaps using coefficients from elastic-net

### G12878

#### 10 kb

#None
##OC
enetcoefs_none_oc_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/ENET/enetModel.rds")
enetcoefs_none_oc_10kb_gm12878 <- coef(enetcoefs_none_oc_10kb_gm12878$finalModel, s=enetcoefs_none_oc_10kb_gm12878$bestTune$lambda)
##OP
enetcoefs_none_op_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/ENET/enetModel.rds")
enetcoefs_none_op_10kb_gm12878 <- coef(enetcoefs_none_op_10kb_gm12878$finalModel, s=enetcoefs_none_op_10kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_none_dist_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/ENET/enetModel.rds")
enetcoefs_none_dist_10kb_gm12878 <- coef(enetcoefs_none_dist_10kb_gm12878$finalModel, s=enetcoefs_none_dist_10kb_gm12878$bestTune$lambda)

#RUS
##OC
enetcoefs_RUS_oc_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/ENET/enetModel.rds")
enetcoefs_RUS_oc_10kb_gm12878 <- coef(enetcoefs_RUS_oc_10kb_gm12878$finalModel, s=enetcoefs_RUS_oc_10kb_gm12878$bestTune$lambda)
##OP
enetcoefs_RUS_op_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/ENET/enetModel.rds")
enetcoefs_RUS_op_10kb_gm12878 <- coef(enetcoefs_RUS_op_10kb_gm12878$finalModel, s=enetcoefs_RUS_op_10kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_RUS_dist_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/ENET/enetModel.rds")
enetcoefs_RUS_dist_10kb_gm12878 <- coef(enetcoefs_RUS_dist_10kb_gm12878$finalModel, s=enetcoefs_RUS_dist_10kb_gm12878$bestTune$lambda)

#ROS
##OC
enetcoefs_ROS_oc_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/ENET/enetModel.rds")
enetcoefs_ROS_oc_10kb_gm12878 <- coef(enetcoefs_ROS_oc_10kb_gm12878$finalModel, s=enetcoefs_ROS_oc_10kb_gm12878$bestTune$lambda)
##OP
enetcoefs_ROS_op_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/ENET/enetModel.rds")
enetcoefs_ROS_op_10kb_gm12878 <- coef(enetcoefs_ROS_op_10kb_gm12878$finalModel, s=enetcoefs_ROS_op_10kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_ROS_dist_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/ENET/enetModel.rds")
enetcoefs_ROS_dist_10kb_gm12878 <- coef(enetcoefs_ROS_dist_10kb_gm12878$finalModel, s=enetcoefs_ROS_dist_10kb_gm12878$bestTune$lambda)

#SMOTE
##OC
enetcoefs_SMOTE_oc_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/ENET/enetModel.rds")
enetcoefs_SMOTE_oc_10kb_gm12878 <- coef(enetcoefs_SMOTE_oc_10kb_gm12878$finalModel, s=enetcoefs_SMOTE_oc_10kb_gm12878$bestTune$lambda)
##OP
enetcoefs_SMOTE_op_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/ENET/enetModel.rds")
enetcoefs_SMOTE_op_10kb_gm12878 <- coef(enetcoefs_SMOTE_op_10kb_gm12878$finalModel, s=enetcoefs_SMOTE_op_10kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_SMOTE_dist_10kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/enetModel.rds")
enetcoefs_SMOTE_dist_10kb_gm12878 <- coef(enetcoefs_SMOTE_dist_10kb_gm12878$finalModel, s=enetcoefs_SMOTE_dist_10kb_gm12878$bestTune$lambda)

enetcoefsmat <- cbind(enetcoefs_none_oc_10kb_gm12878[-1,1], 
                      enetcoefs_ROS_oc_10kb_gm12878[-1,1],
                      enetcoefs_RUS_oc_10kb_gm12878[-1,1], 
                      enetcoefs_SMOTE_oc_10kb_gm12878[-1,1],

                      enetcoefs_none_op_10kb_gm12878[-1,1], 
                      enetcoefs_ROS_op_10kb_gm12878[-1,1],
                      enetcoefs_RUS_op_10kb_gm12878[-1,1], 
                      enetcoefs_SMOTE_op_10kb_gm12878[-1,1],

                      enetcoefs_none_dist_10kb_gm12878[-1,1],
                      enetcoefs_ROS_dist_10kb_gm12878[-1,1],
                      enetcoefs_RUS_dist_10kb_gm12878[-1,1],
                      enetcoefs_SMOTE_dist_10kb_gm12878[-1,1])

enetcoefsmat[enetcoefsmat == 0] <- NA

#colnames(enetcoefsmat) <- c(paste(c("OC","OP","Distance"), rep("None", 3), sep=" "),
#paste(c("OC","OP","Distance"), rep("RUS", 3), sep=" "),
#paste(c("OC","OP","Distance"), rep("ROS", 3), sep=" "),
#paste(c("OC","OP","Distance"), rep("SMOTE", 3), sep=" "))
colnames(enetcoefsmat) <- rep(c("None", "ROS", "RUS", "SMOTE"), 3)
rownames(enetcoefsmat) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", rownames(enetcoefsmat))

my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(10))

##### Overlap Counts

```{r}
heatmap.2(enetcoefsmat[,1:4],
          #cellnote = cormat,  # same data set for cell labels
          #main = "Correlation", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          dendrogram="row",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.75,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          sepwidth=c(0.001,0.001),
          sepcolor="black",
          colsep=1:4,
          rowsep=1:nrow(enetcoefsmat),
          keysize = 1) 
```

#####Overlap Percents

```{r}
heatmap.2(enetcoefsmat[,5:8],
          #cellnote = cormat,  # same data set for cell labels
          #main = "Correlation", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          dendrogram="row",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.75,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          sepwidth=c(0.001,0.001),
          sepcolor="black",
          colsep=1:4,
          rowsep=1:nrow(enetcoefsmat),
          keysize = 1) 
```

##### Distances

```{r}
heatmap.2(enetcoefsmat[,9:12],
          #cellnote = cormat,  # same data set for cell labels
          #main = "Correlation", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          dendrogram="row",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.75,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          sepwidth=c(0.001,0.001),
          sepcolor="black",
          colsep=1:4,
          rowsep=1:nrow(enetcoefsmat),
          keysize = 1) 
```

#### 25kb

#None
##OC
enetcoefs_none_oc_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/ENET/enetModel.rds")
enetcoefs_none_oc_25kb_gm12878 <- coef(enetcoefs_none_oc_25kb_gm12878$finalModel, s=enetcoefs_none_oc_25kb_gm12878$bestTune$lambda)
##OP
enetcoefs_none_op_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/ENET/enetModel.rds")
enetcoefs_none_op_25kb_gm12878 <- coef(enetcoefs_none_op_25kb_gm12878$finalModel, s=enetcoefs_none_op_25kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_none_dist_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/ENET/enetModel.rds")
enetcoefs_none_dist_25kb_gm12878 <- coef(enetcoefs_none_dist_25kb_gm12878$finalModel, s=enetcoefs_none_dist_25kb_gm12878$bestTune$lambda)

#RUS
##OC
enetcoefs_RUS_oc_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/ENET/enetModel.rds")
enetcoefs_RUS_oc_25kb_gm12878 <- coef(enetcoefs_RUS_oc_25kb_gm12878$finalModel, s=enetcoefs_RUS_oc_25kb_gm12878$bestTune$lambda)
##OP
enetcoefs_RUS_op_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/ENET/enetModel.rds")
enetcoefs_RUS_op_25kb_gm12878 <- coef(enetcoefs_RUS_op_25kb_gm12878$finalModel, s=enetcoefs_RUS_op_25kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_RUS_dist_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/ENET/enetModel.rds")
enetcoefs_RUS_dist_25kb_gm12878 <- coef(enetcoefs_RUS_dist_25kb_gm12878$finalModel, s=enetcoefs_RUS_dist_25kb_gm12878$bestTune$lambda)

#ROS
##OC
enetcoefs_ROS_oc_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OC/ENET/enetModel.rds")
enetcoefs_ROS_oc_25kb_gm12878 <- coef(enetcoefs_ROS_oc_25kb_gm12878$finalModel, s=enetcoefs_ROS_oc_25kb_gm12878$bestTune$lambda)
##OP
enetcoefs_ROS_op_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OP/ENET/enetModel.rds")
enetcoefs_ROS_op_25kb_gm12878 <- coef(enetcoefs_ROS_op_25kb_gm12878$finalModel, s=enetcoefs_ROS_op_25kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_ROS_dist_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/ENET/enetModel.rds")
enetcoefs_ROS_dist_25kb_gm12878 <- coef(enetcoefs_ROS_dist_25kb_gm12878$finalModel, s=enetcoefs_ROS_dist_25kb_gm12878$bestTune$lambda)

#SMOTE
##OC
enetcoefs_SMOTE_oc_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OC/ENET/enetModel.rds")
enetcoefs_SMOTE_oc_25kb_gm12878 <- coef(enetcoefs_SMOTE_oc_25kb_gm12878$finalModel, s=enetcoefs_SMOTE_oc_25kb_gm12878$bestTune$lambda)
##OP
enetcoefs_SMOTE_op_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OP/ENET/enetModel.rds")
enetcoefs_SMOTE_op_25kb_gm12878 <- coef(enetcoefs_SMOTE_op_25kb_gm12878$finalModel, s=enetcoefs_SMOTE_op_25kb_gm12878$bestTune$lambda)
##Distance
enetcoefs_SMOTE_dist_25kb_gm12878 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/ENET/enetModel.rds")
enetcoefs_SMOTE_dist_25kb_gm12878 <- coef(enetcoefs_SMOTE_dist_25kb_gm12878$finalModel, s=enetcoefs_SMOTE_dist_25kb_gm12878$bestTune$lambda)

enetcoefsmat <- cbind(enetcoefs_none_oc_25kb_gm12878[-1,1], 
                      enetcoefs_ROS_oc_25kb_gm12878[-1,1],
                      enetcoefs_RUS_oc_25kb_gm12878[-1,1], 
                      enetcoefs_SMOTE_oc_25kb_gm12878[-1,1],

                      enetcoefs_none_op_25kb_gm12878[-1,1], 
                      enetcoefs_ROS_op_25kb_gm12878[-1,1],
                      enetcoefs_RUS_op_25kb_gm12878[-1,1], 
                      enetcoefs_SMOTE_op_25kb_gm12878[-1,1],

                      enetcoefs_none_dist_25kb_gm12878[-1,1],
                      enetcoefs_ROS_dist_25kb_gm12878[-1,1],
                      enetcoefs_RUS_dist_25kb_gm12878[-1,1],
                      enetcoefs_SMOTE_dist_25kb_gm12878[-1,1])

enetcoefsmat[enetcoefsmat == 0] <- NA

#colnames(enetcoefsmat) <- c(paste(c("OC","OP","Distance"), rep("None", 3), sep=" "),
#paste(c("OC","OP","Distance"), rep("RUS", 3), sep=" "),
#paste(c("OC","OP","Distance"), rep("ROS", 3), sep=" "),
#paste(c("OC","OP","Distance"), rep("SMOTE", 3), sep=" "))
colnames(enetcoefsmat) <- rep(c("None", "ROS", "RUS", "SMOTE"), 3)
rownames(enetcoefsmat) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", rownames(enetcoefsmat))

my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(10))

##### Overlap Counts

```{r}
heatmap.2(enetcoefsmat[,1:4],
          #cellnote = cormat,  # same data set for cell labels
          #main = "Correlation", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          dendrogram="row",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.75,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          sepwidth=c(0.001,0.001),
          sepcolor="black",
          colsep=1:4,
          rowsep=1:nrow(enetcoefsmat),
          keysize = 1) 
```

#####Overlap Percents

```{r}
heatmap.2(enetcoefsmat[,5:8],
          #cellnote = cormat,  # same data set for cell labels
          #main = "Correlation", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          dendrogram="row",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.75,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          sepwidth=c(0.001,0.001),
          sepcolor="black",
          colsep=1:4,
          rowsep=1:nrow(enetcoefsmat),
          keysize = 1) 
```

##### Distances

```{r}
my_palette2 <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(10))
my_palette2 <- c(my_palette2[1:20], "gray", my_palette2[21:39])
my_palette2 <- colorRampPalette(c("blue", "white", "red"))(n=100)

color.palette  <- c(colorRampPalette(c("blue", "gray", "red"))(n=99)[1:49], 
                    "#F8F8F8", 
                    colorRampPalette(c("yellow", "orange", "red"))(n=99)[50:99])
enetcoefsmat2 <- enetcoefsmat
enetcoefsmat2[,9:12][is.na(enetcoefsmat2[,9:12])] <- 0

col_breaks = c(seq(-.999,-.900,length=2),
           seq(-.899,-.800,length=2),
           seq(-.799,-.700,length=2),
           seq(-.699,-.600,length=2),
           seq(-.599,-.500,length=2),
           seq(-.499,-.400,length=2),
           seq(-.399,-.300,length=2),
           seq(-.299,-.200,length=2),
           seq(-.199,-.100,length=2),
           seq(-.099,-.001,length=2),
           seq(0,0.1,length=1),  # for gray
           seq(.001,.099,length=2),
           seq(.100,.199,length=2),
           seq(.200,.299,length=2),
           seq(.300,.399,length=2),
           seq(.400,.499,length=2),
           seq(.500,.599,length=2),
           seq(.600,.699,length=2),
           seq(.700,.799,length=2),
           seq(.800,.899,length=2),
           seq(.900,.999,length=2))

col_breaks = c(seq(-1,-.7,length=20),
           seq(-.699,-.001,length=20),
           seq(0,0.001,length=1),  # for white
           seq(.001,.699,length=20),
           seq(.7,1,length=20))

heatmap.2(enetcoefsmat2[,9:12],
          #cellnote = cormat,  # same data set for cell labels
          #main = "Correlation", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins =c(5,5),     # widens margins around plot
          col=my_palette2,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          dendrogram="row",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.75,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          sepwidth=c(0.001,0.001),
          sepcolor="black",
          colsep=1:4,
          rowsep=1:nrow(enetcoefsmat),
          keysize = 1) 
```
